---
title: "Midterm-1-Review"
format: html
editor: visual
---
Here are the revised questions that make sense within the context of R programming and the use of `dplyr` and `ggplot2` in data analysis:

1. **Variable Types in `mtcars`**:
Identify and specify the variable types (e.g., numeric, factor) in the `mtcars` dataset. Use the `?mtcars` command to access the help page and examine the variables more closely.

2. **Mean Horsepower by Transmission**:
Calculate the mean horsepower (`hp`) grouped by transmission type (`am`). Ensure to differentiate between automatic and manual transmissions.

3. **Categorize `mpg` Using `case_when`**:
Categorize the `mpg` variable into four levels ("Low", "Moderate", "High", "Very High") using the `case_when` function. Define the levels based on the quartiles (Q1, Median/Q2, and Q3). Name the new variable as `split_mpg`.

4. **Violin Plot of Horsepower by `split_mpg`**:
Create a violin plot to visualize the distribution of horsepower (`hp`) across the four `split_mpg` categories you defined in the previous question.

5. **Highest Horsepower Group in `split_mpg`**:
Determine which `split_mpg` group has the highest average horsepower. Summarize the results in a way that clearly identifies the group with the highest horsepower.

6. **Comparing Weight and MPG Visualizations**:
Produce two different visualizations to display the relationship between weight (`wt`) and miles per gallon (`mpg`), while also showing the different types of transmission (`am`). After creating the plots, provide an analysis of which visualization is more effective in illustrating the differences between transmission types and why.

7. **Simulate Student Data with `rnorm`**:
Use the `rnorm` function to simulate three variables representing a sample of 100 students: GPA (mean: 3.5, standard deviation: 0.1), study hours (mean: 35, standard deviation: 3), and work hours (mean: 15, standard deviation: 1).

8. **Study Hours and Work Hours Relationship Plot**:
Based on the simulated data from the previous question, create a plot that visualizes the relationship between study hours and work hours. Analyze the plot to understand the distribution and correlation, if any, between the two variables.



9. **Evaluating a Misleading Visualization in `ggplot2`**:
Review the following plot. Explain why representing months as numerical values on the x-axis could lead to a misleading interpretation of the `Ozone` data in the `airquality` dataset.

```{r}
library(ggplot2)

ggplot(airquality, aes(x = Month, y = Ozone)) + 
  geom_point() 
```

10. **Improving the Month Representation in a ggplot**:
The `airquality` dataset uses numeric values for months. Modify the ggplot below to replace numeric month values with their corresponding month names for clearer interpretation. Use the `airquality` dataset and display the `Ozone` levels over the months.
```{r}
## Answer for question 10.
```

11. **Debugging Line Display in ggplot2**:
The following ggplot code is intended to show the relationship between `Solar.R` and `Ozone`, with lines representing different months. However, the lines are not displaying as expected. Correct the code so that each month's data is connected by lines with distinct line types.
```{r}
ggplot(airquality, aes(x = Solar.R, y = Ozone)) + 
  geom_point(aes(color = factor(Month))) +
  geom_line(aes(linetype = factor(Month, levels = c("Five","Six","Seven","Eight","Nine"))))
```
Provide the corrected ggplot code to ensure that lines are properly displayed for each month, and explain why the original code failed to show the lines.
