---
title: "PA 2: Using Data Visualization to Find the Penguins"
format: html
self-contained: true
execute: 
  echo: true
  error: true
  code-fold: true
---

<!-- Insert your setup code chunk here. This is just good practice. -->

[Download the .qmd template](PA2-ggplot-palmer-penguins.qmd) and save it in a reasonable location.

Today you will be exploring different types of visualizations to uncover which species of penguins reside on different islands.

Some advice:

+ Work with those around you

+ If you aren't sure how to make a specific plot or how to customize a plot, look over the class text for this week and make use of the [R graphics cheat sheets](https://earobinson95.github.io/stat331-calpoly-text/02-tidy-data-and-basics-of-graphics.html#rgraphics).

+ Google is your friend! If you aren't sure how to accomplish a certain task, type what you are trying to accomplish into Google and see what other people are saying. Generally, adding `ggplot` to the end of your search will help make your search results more relevant!

## Getting Started

We will be creating visualizations using the `ggplot2` package.

For this activity, we will be exploring the `penguins` data from the `palmerpenguins` package, which has [fantastic documentation with really awesome artwork](https://allisonhorst.github.io/palmerpenguins/). So, you will need to install the **palmerpenguins** package.

`install.packages("palmerpenguins")`

## Creating a Set-up Code Chunk

1.  Insert a code chunk at the beginning of your document (directly under the YAML).
2.  Name the code chunk *setup*.
3.  Specify the code chunk option with the hashpipe `#|` that the messages from loading in the packages should not be included.
4.  Load in the `tidyverse` or `ggplot2` package.
5.  Load in the `palmerpenguins` package.

::: callout-note
Naming your code chunk "setup" has special properties in .qmd - specifically this code chunk will run automatically when you try to run a subsequent code chunk. This ensures all packages and any other specifications for your document are loaded and will not cause you errors or messages.
:::

## Data set `penguins`

I like to tangibly see the data set I will be working with so I am going to pull the `penguins` data into my R environment.

```{r}
data(penguins)
```

6. Get to know your data, what are the variables and what units are they measured in? What does each row represent?

## Graphics

::: callout-note
Make sure to give your plots reader friendly axes labels!
:::

7. Use `ggplot2` to draw a barchart of `species`. Map `species` to the fill color of the barchart.

```{r}

```

8. Create a scatterplot of the relationship between the bill length (`bill_length_mm`) and bill depth (`bill_depth_mm`).

```{r}

```

9. Add an aesthetic to differentiate the species by color.

```{r}

```
10. Add the location of the penguins, `island`, to your visualization. There may be more than one method to address this, however, one method will more easily address the questions below.

```{r}

```

## Canvas Quiz

::: callout-tip
### Use the plots you created to address the following questions on Canvas:

1.  Which species of penguins is represented **least** in the Palmer Penguins data set?

2.  Which species of penguins are found on Biscoe Island?

3.  Which species of penguins are found on Dream Island?

4.  Which species of penguins are found on Torgersen Island?

5.  Color deficiencies can make differentiating groups difficult - which aesthetic could you add to your plot to double encode species?
:::

### Useful References

 + [Customizing ggplot2 graphics](http://r-statistics.co/Complete-Ggplot2-Tutorial-Part2-Customizing-Theme-With-R-Code.html)
 + [A group-created slideshow of awful graphics](https://docs.google.com/presentation/d/1ixRsGI8BWgT2zA5AbB61eMLUbJgObvBCZt-OnZDzvX4/edit#slide=id.p)


